---
title: "Activity Monitoring"
author: "Olga Cuznetova"
date: "December 7, 2016"
output: md_document
        
---

######Loading data:
```{r}
#setwd("ReproducibleResearch/")
activity <- read.csv("activity.csv")
```

###What is mean total number of steps taken per day?

 - Total number of steps taken per day: 
 
```{r}
totalSteps <- aggregate(steps ~ date, activity, sum)
head(totalSteps)
```

 - Histogram of the total number of steps taken each day:
 
```{r}
hist(totalSteps$steps, main = paste("Total Steps Taken per Day"), col="green", xlab="Number of Steps", ylab="Frequency")
```

 - Mean number of steps
 
```{r}
meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
meanSteps
``` 
 
 - Median number of steps 
 
```{r}
medianSteps <- median(totalSteps$steps, na.rm = TRUE)
medianSteps
``` 
 
 
###What is the average daily activity pattern?
 
```{r}
stepsInterval <- aggregate(steps ~ interval, activity, mean)
head(stepsInterval)
```  
 
 - Plot of Average number of steps by interval: 
 
```{r}
plot(stepsInterval$interval, stepsInterval$steps, type="l", col="green", xlab="5-minute Intervals", ylab="Average Number of Steps Taken",main="Average Number of Steps by Interval")
```  
 
 - Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
 
```{r}
intervalWMaxSteps <- stepsInterval[which.max(stepsInterval$steps),1]
intervalWMaxSteps
plot(stepsInterval$interval, stepsInterval$steps, type="l", col="green", xlab="5-minute Intervals", ylab="Average Number of Steps Taken",main="Average Number of Steps by Interval")
abline(v = 835, col = "orange")
```  
 
 
###Imputing missig values:

 - Number of rows with missing values:
 
```{r}
incompleteActivity <- sum(is.na(activity))
incompleteActivity 
```  
 
 - All the missing values were filled in with mean number of steps
 
```{r}
imputedActivity <- activity
nas <- is.na(imputedActivity$steps)
meanInterval <- tapply(imputedActivity$steps, imputedActivity$interval, mean, na.rm=TRUE, simplify=TRUE)
imputedActivity$steps[nas] <- meanInterval[as.character(imputedActivity$interval[nas])]
``` 

 -  Histogram of Total steps taken per day using imputed data
 
```{r}
totalStepsImputed <- aggregate(steps ~ date, imputedActivity, sum)
hist(totalStepsImputed$steps, main = paste("Total Steps Taken per Day"), col="orange", xlab="Number of Steps", ylab="Frequency")

``` 

 - Histogram that show difference between imputed and non-imputed data
 
```{r}
hist(totalStepsImputed$steps, main = paste("Total Steps Taken per Day"), col="orange", xlab="Number of Steps", ylab="Frequency")
hist(totalSteps$steps, main = paste("Total Steps Taken per Day"), col="green", xlab="Number of Steps", ylab="Frequency", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("orange", "green"), lwd=10)
```  
 
 - Mean, meadian using imputed data
 
```{r}
meanStepsImputed <- mean(totalStepsImputed$steps)
medianStepsImputed <- median(totalStepsImputed$steps)
meanStepsImputed
medianStepsImputed
```  
 
 - The following difference is observed between imputed and non-imputed data:
 
```{r}
meanD <- meanStepsImputed - meanSteps
medianD <- medianStepsImputed - medianSteps
totalD <- sum(totalStepsImputed$steps) - sum(totalSteps$steps)
meanD
medianD
totalD
```   
 
###Are there differences in activity patterns between weekdays and weekends? 
 
 - Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day:
 
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
imputedActivity$day = as.factor(ifelse(is.element(weekdays(as.Date(imputedActivity$date)),weekdays), "Weekday", "Weekend"))
```    
 
 - Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis):
 
```{r}
stepsInterval <- aggregate(steps ~ interval + day, imputedActivity, mean)

library(lattice)

xyplot(stepsInterval$steps ~ stepsInterval$interval|stepsInterval$day, main="Average Number of Steps by Interval",xlab="5-minute Intervals", ylab="Average Number of Steps Taken",layout=c(1,2), type="l")
```   
